The Calculus of Linear Constructions

### 2021-10-19 writing/lics22/lics22.tex
lics22

\renewcommand{\syntleft}{}
\renewcommand{\syntright}{}
\setlength{\grammarparsep}{7pt}
\setlength{\grammarindent}{7em}
\newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
\newcommand{\rname}[1]{\textsc{\footnotesize #1}}
\newcommand{\pure}[1]{|#1|}
\newcommand{\sort}{\textbf{sort}}
\newcommand{\type}{\textbf{type}}
\newcommand{\refl}{\text{refl}}
\newcommand{\letin}[3]{\ $\text{let }#1\text{ := }#2\text{ in }#3$\ }
\newcommand{\ind}{\ $\text{ind}$\ }
\newcommand{\iter}{\ $\text{iter}$\ }
\newcommand{\unit}{\text{unit}}
\newcommand{\new}{\textbf{new}}
\newcommand{\free}{\textbf{free}}
\newcommand{\get}{\textbf{get}}
\newcommand{\set}{\textbf{set}}
\newcommand{\utype}{:_{\scriptscriptstyle U}}
\newcommand{\ltype}{:_{\scriptscriptstyle L}}
\newcommand{\stype}[1]{:_#1}
\newcommand{\step}{\leadsto}
\newcommand{\pstep}{\leadsto}
\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
\newcommand{\erase}[1]{\llbracket #1 \rrbracket}

@@

  We prove the following lemma to commute erasure and substitution whenever needed.
  \begin{lemma} 
    Erasure commutes with substitution. For any CLC terms $m$, $n$, and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
  \end{lemma}

  For the following lemma, we refer to the reductions in CLC as $\pstep_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\pstep_{\scriptscriptstyle \text{CC$\omega$}}$.
  \begin{lemma} \label{preserve} 
    Erasure preserves reduction. For any CLC terms $m$, and $n$, if there is $m \pstep_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \pstep_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
  \end{lemma}

  \begin{lemma} \label{embedding} 
    Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
  \end{lemma}

  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of lemmas \ref{preserve}, and \ref{embedding}. This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.

@@

>>>>

\renewcommand{\syntleft}{}
\renewcommand{\syntright}{}
\setlength{\grammarparsep}{7pt}
\setlength{\grammarindent}{7em}
\newcommand{\indalt}[1][2]{\\\hspace*{-1.2em}\textbar\quad}
\newcommand{\rname}[1]{\textsc{\footnotesize #1}}
\newcommand{\pure}[1]{|#1|}
\newcommand{\sort}{\textbf{sort}}
\newcommand{\type}{\textbf{type}}
\newcommand{\refl}{\text{refl}}
\newcommand{\letin}[3]{\ $\text{let }#1\text{ := }#2\text{ in }#3$\ }
\newcommand{\ind}{\ $\text{ind}$\ }
\newcommand{\iter}{\ $\text{iter}$\ }
\newcommand{\unit}{\text{unit}}
\newcommand{\new}{\textbf{new}}
\newcommand{\free}{\textbf{free}}
\newcommand{\get}{\textbf{get}}
\newcommand{\set}{\textbf{set}}
\newcommand{\utype}{:_{\scriptscriptstyle U}}
\newcommand{\ltype}{:_{\scriptscriptstyle L}}
\newcommand{\stype}[1]{:_#1}
\newcommand{\step}{\leadsto}
\newcommand{\pstep}{\leadsto}
\newcommand{\mrg}[3]{#1\ddagger#2\ddagger#3}
\newcommand{\erase}[1]{\llbracket #1 \rrbracket}
\newcommand{\lift}[1]{\llparenthesis #1 \rrparenthesis}

@@

  We prove the following lemma to commute erasure and substitution whenever needed.
  \begin{lemma} 
    Erasure commutes with substitution. For any CLC terms $m$, $n$ and some variable $x$, $\erase{m[n/x]} = \erase{m}[\erase{n}/x]$. 
  \end{lemma}

  For the following lemma, we refer to the reductions in CLC as $\step_{\scriptscriptstyle \text{CLC}}$, and the reductions in CC$\omega$ as $\step_{\scriptscriptstyle \text{CC$\omega$}}$.
  \begin{theorem} \label{preserve} 
    Erasure preserves reduction. For any CLC terms $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CLC}} n$, then there is $\erase{m} \step_{\scriptscriptstyle \text{CC$\omega$}} \erase{n}$.
  \end{theorem}

  \begin{theorem} \label{embedding} 
    Embedding. If $\Gamma \vdash m \stype{s} A$ is a valid judgment in CLC, then $\erase{\Gamma} \vdash \erase{m} : \erase{A}$ is a valid judgment in CC$\omega$.
  \end{theorem}

  If there exists some well typed CLC term with an infinite sequence of reductions, erasure will embed this term into a well typed CC$\omega$ term along with its infinite sequence of reductions by virtue of theorems \ref{preserve}, and \ref{embedding}. This is contradictory to the strong normalization property of CC$\omega$ proven through the Girard-Tait method \cite{ecc}, so this hypothetical term does not exist in CLC.

@@

  \subsubsection{Backwards Capability}
  To show that CLC is backwards compatible with CC$\omega$, we construct a function that annotates CC$\omega$ terms with sort $U$, lifting them into the non-linear fragment of CLC.

  \begin{definition}
    \begin{align*}
      \lift{x} &= x \\
      \lift{Prop} &= U_* \\
      \lift{Type_k} &= U_k \\
      \lift{(x : A) \rightarrow B} &= (x \utype \lift{A}) \rightarrow \lift{B} \\
      \lift{\lambda x.n} &= \lambda x.\lift{n} \\
      \lift{m\ n} &= \lift{m}\ \lift{n}
    \end{align*}
  \end{definition}

  With slight overloading of notation, we define lifting for CC$\omega$ recursively.
  \begin{definition}
    \begin{align*}
      \lift{\epsilon} &= \epsilon \\
      \lift{\Gamma, x : A} &= \lift{\Gamma}, x \utype \lift{A}
    \end{align*}
  \end{definition}

  The following lemmas and theorems are all proved following a similar procedure to proving the soundness of erasure.
  \begin{lemma}
    Lift commutes with substitution. For any CC$\omega$ terms $m$, $n$ and some variable $x$, $\lift{m[n/x]} = \lift{m}[\lift{n}/x]$.
  \end{lemma}

  \begin{theorem}
    Lift preserves reduction. For any CC$\omega$ term $m$ and $n$, if there is $m \step_{\scriptscriptstyle \text{CC}\omega} n$, then there is $\lift{m} \step_{\scriptscriptstyle \text{CLC}} \lift{n}$.
  \end{theorem}

  \begin{theorem}
    Lifting. If $\Gamma \vdash m : A$ is a valid judgment in CC$\omega$, then $\lift{\Gamma} \vdash \lift{m} \utype \lift{A}$ is a valid judgment in CLC.
  \end{theorem}
